Grade Retention and Mixed Race

### 2022-09-09 products/paper/main.qmd
Update text on results for new figures

Controlling for resource variables has some effect on the placement of part-Black biracial respondents. In the case of Black/White biracial respondents, Black/White students initially have a risk of BBEG slightly higher than and statistically distinguishable from White students. After controlling for cultural and family resources, however, Black/White respondents have a risk of BBEG that is slightly lower and statistically distinguishable from White students. This change reflects the relatively low level of parental education and income among Black/White students, which is much closer to Black students than White students, as shown on @tbl-descriptive.

An opposite pattern is observed for Black/Latino respondents. Black/Latino respondents initially have a risk of grade retention substantially lower than both Black and Latino students, who have similar risks. However, after controlling for cultural and family resources, Black/Latino students have a risk of BBEG that is indistinguishable from the risk of Latino students and substantially lower than Black students. The change largely reflects the fact that Black/Latino students have greater resources than Latino (but not Black) students and once this advantage is held constant, they have a similar risk.

The role of parental resources is harder to determine for Black/Asian and Black/Indigenous students due to the greater uncertainty in these estimates. Controlling for resources moves the point estimate for Black/Asian students from a point estimate close to Asian students to more of a halfway position between the two constituent groups. Because Black/Indigenous students had resources halfway between their constituent groups, controlling for these resources hardly affects the relative placement of Black/Indigenous students at all.

@fig-other-models shows the results for the remaining biracial groups. In general, the results here show that biracial students are more likely to have a risk of BBEG more like the monoracial constituent group with the lower risk than the higher risk, with some important caveats discussed below.

@@

For all groups, the full models generally indicate that each biracial group has a risk of BBEG similar to the risk of the lower monoracial constituent group or has risk higher than this group but less than the halfway point. There are two exceptions to this general pattern. First, the point estimate of the probability of BBEG for the Indigenous/Latino group is roughly halfway between the two constituent groups and statistically distinguishable from both. Second, Latino/Asian students are the only case where the point estimate of the risk for the biracial group is higher than either monoracial group. However, the confidence bands suggest that the risk for this group cannot be statistically differentiated from either of the two monoracial constituent groups. This result only emerges in the model that controls for resources, Latino/Asian students come from more advantaged households than both Latino and Asian students, on average.

Aside from the effect on Latino/Asian students, controlling for resources has moderate effects on the placement of the remaining biracial groups. White/Asian students are the only case with a substantial shift, owing to their highly advantaged households. After controlling for resources, the risks of BBEG for White/Asian students are indistinguishable from the risk of Asian students, but still substantially and statistically distinguishable from the higher risk to White students.

>>>>

Controlling for resource variables has some effect on the placement of part-Black biracial respondents. In the case of Black/White biracial respondents, Black/White students initially have a risk of BBEG slightly higher than and statistically distinguishable from White students. After controlling for cultural and family resources, however, Black/White respondents have a risk of BBEG that is slightly lower but not statistically distinguishable from White students. This change reflects the relatively low level of parental education and income among Black/White students, which is much closer to Black students than White students, as shown on @tbl-descriptive.

Black/Latino students start with a very different position, having a risk of BBEG substantially lower than both Black and Latino students, who have sonewgat similar risks. However, after controlling for cultural and family resources, Black/Latino students have a risk of BBEG that is statistically indistinguishable from the risk of Latino students and substantially lower than and statistically distinguishable from Black students. The change largely reflects the fact that Black/Latino students have greater resources than Latino (but not Black) students and once this advantage is held constant, they have a similar risk. After controlling for family resources, both Black/White and Black/Latino students have a risk of BBEG similar to their monoracial constituent group with lower risk.

The role of parental resources is harder to determine for Black/Asian and Black/Indigenous students due to the greater uncertainty in these estimates. Controlling for resources moves the point estimate for Black/Asian students from a point estimate close to Asian students to more of a halfway position between the two constituent groups. Black/Indigenous students have BBEG risk roughly halfway between their two constituent groups in both models. However, this finding of a halfway position for both groups is highly tentative owing to the wide confidence intervals around both groups, due to small sample size.

@fig-other-models shows the results for the remaining biracial groups. In general, the results here show that biracial students are more likely to have a risk of BBEG closer to the monoracial constituent group with the lower risk than the higher risk, with some important caveats discussed below.

@@

For all groups, the full models generally indicate that each biracial group has a risk of BBEG similar to the risk of the lower monoracial constituent group or has risk higher than this group but less than the halfway point. There are two exceptions to this general pattern. First, the point estimate of the probability of BBEG for the Indigenous/Latino group is roughly halfway between the two constituent groups, although it is not statistically distinguishable from the lower group. Second, Latino/Asian students are the only case where the point estimate of the risk for the biracial group is higher than either monoracial group. However, the confidence bands suggest that the risk for this group cannot be statistically differentiated from either of the two monoracial constituent groups who have very similar risks. This result only emerges in the model that controls for resources, because Latino/Asian students come from more advantaged households than both Latino and Asian students, on average.

Aside from the effect on Latino/Asian students, controlling for resources has moderate effects on the placement of the remaining biracial groups. White/Asian students are the only case with a substantial shift, owing to their highly advantaged households. Initially white/Asian students have risks of BBEG substantially below both White and Asian students. After controlling for resources, the risks of BBEG for White/Asian students are indistinguishable from the those for Asian students, but still substantially and statistically distinguishable from the higher risk to White students.
